---
title: Code optimization using commuting example
output: pdf
---

### Commuting rate formula:

The expected commuting rate $\sigma_{ji}$ from a source node $j$ and a destination node $i$ is:

\begin{equation}
\langle \sigma_{ji} \rangle = \sigma_j \frac{n_j n_i}{(n_j + s_{ji})(n_j + n_i + s_{ji})}
\end{equation}

where $n_j$ and $n_i$ are the populations for nodes $j$ and $i$, and $s_{ji}$ is the total population (excluding $j$ and $i$) in a circle centered on $j$ with a radius equal to the distance between $j$ and $i$. The total commuting rate of individuals in $j$ is $\sigma_j = N_c/N$ = 11%, where $N_c$ is the total number of commuters and $N$ is the total population in the country. 

\  

### External libraries used in this document:
`here` and `igraph`


### Sourcing code:

```{r}
invisible(lapply(here::here(list.files("R", full = TRUE)), source))
```

We're specifically working with code in the `R/commuting.R` file. 

### Reading in example data:

```{r}
# read in data and make a subset of it
g = readRDS(here::here("inst/sampleData/flu-g.RDS"))
g = igraph::induced.subgraph(g, c("890", sample(1:1000, 250)))

# network edges (distances between nodes)
head(igraph::as_data_frame(g, "edges"))

# network node information
head(igraph::as_data_frame(g, "vertices"))
```


### Calculating commuting proportions:
```{r, results = "hide"}
# calculate commuting rates over it
t1 = system.time(disnet_commuting(g))
t2 = system.time(disnet_commuting2(g))
```

### Time taken by the two methods:
```{r}
# Method 1
t1

# Method 2
t2
```

There's some improvement with the second method but definitely scope for better code optimization! 
